---
title: "The Impact of the Affordable Care Act on Healthcare Access in Minnesota Counties"
subtitle: "A County-Level Analysis of Changes in Coverage Before and After ACA Implementation"
authors: "Jacob Posner, Evan Burns, Max Clifford"
date: "December, 2024"
format: 
  html:
    embed-resources: true
    toc: true
    theme: superhero
---

```{r include = FALSE}
knitr::opts_chunk$set(
  warning = FALSE,
  message = FALSE,
  collapse = TRUE, 
  echo = FALSE, 
  fig.height = 5, 
  fig.width = 7,
  fig.align = 'center')
```

```{r}
# load packages
library(dplyr)
library(tidyverse)
library(ggplot2)
library(dagitty)
library(readr)
library(stringr)
library(viridis)
```
```{r}
# Load Data

```

## Introduction

    On March 23, 2010, President Obama signed the Affordable Care Act (ACA) into law, a landmark reform aimed at making healthcare more accessible and affordable for all Americans. Prior to the ACA, the uninsured rate in the U.S. was approximately 15% (The Share of Americans Without Health Insurance in 2023 Remained Low, 2024), with many families facing significant barriers to obtaining coverage. Employer-sponsored insurance was the primary source of healthcare for many individuals and families, but not all employers provided it, leading to gaps in coverage. Additionally, pre-existing condition exclusions, high out-of-pocket costs, and restrictive Medicaid eligibility prevented many people, particularly those in lower income brackets, young adults, and rural populations, from obtaining adequate insurance.

    The ACA aimed to address gaps in healthcare access by expanding Medicaid, prohibiting discrimination based on pre-existing conditions, and ensuring that insurance plans covered essential health benefits. While the ACA has significantly increased healthcare access, questions remain about its long-term effects, particularly across different income levels, age groups, and county sizes. Our research seeks to explore how healthcare access has varied across Minnesota’s counties before and after the ACA’s implementation, with a specific focus on differences between large and small counties. This comparison will provide further insights into how the ACA’s impact might differ depending on county characteristics, such as population size and demographics. For an overview of this topic, please visit [MNSure](https://www.mnsure.org/learn-more/aca/index.jsp) or [MN House Research](https://www.house.mn.gov/hrd/pubs/ss/ssmncare.pdf). 

    To investigate these questions, we will examine: *How has access to healthcare varied across Minnesota’s counties before and after the implementation of the ACA?* and *What are the differences in before/after outcomes for large vs. small counties?* These questions are crucial as they explore not only the overall impact of the ACA but also the varying effects based on county size. Understanding these differences can help policymakers refine healthcare strategies, particularly for rural and smaller counties that may face unique challenges.

## Data and Methods

    We used data from the Small Area Health Insurance Estimates (SAHIE) Program through the United States Census Bureau website. SAHIE provides the only source of single-year estimates of health insurance coverage status for all U.S. counties, with breakdowns by economic and demographic characteristics such as age, sex, race, and income (U.S. Census Bureau, 2024). It is collected through annual surveys. For this analysis, we focused on data for Minnesota counties only, as our study is geographically restricted to that state. Key variables include county name (the name of the county), health insurance coverage status (percentage of individuals with and without health insurance), and other demographic characteristics (age, sex, race, and income categories). To prepare the data for analysis, we filtered it to include only Minnesota counties. However, we encountered an issue when trying to analyze race-related data, as it was only available at the state level nationwide, while our analysis required county-level data. To ensure the data set was usable for analysis, we converted all variables to a numeric format. These adjustments allowed us to effectively analyze our data and answer our questions. 

    Paragraph about burns data

    We created a Shiny app to analyze the SAHIE data before and after the ACA's implementation. The app features an interactive map of Minnesota, displaying county-level data on healthcare coverage. Users can select a year from before or after the ACA was implemented to compare healthcare coverage in different counties. The app allows users to choose demographic categories such as age, sex, and income, as well as select variables related to insurance status. These variables represent the percentage of individuals who are insured or uninsured within specific demographic groups, both for selected income categories and across all income levels. The app outputs two maps side by side, comparing the percentage of insured or uninsured individuals in each county for the two selected years. We also used Plotly to enhance the interactivity of the maps, allowing users to zoom in, hover over counties for more detailed information, and compare the data between the selected years. This visualization provides a clear, interactive way to explore how healthcare access has changed across Minnesota’s counties before and after the ACA.

## Results and Analysis

    First Question. Need to include topic, viz, results

    Second question. Need to include topic, viz, results

## Limitations and Next Steps

    

## Conclusion
    

## Works Cited

## Appendix