---
title: "FP2 Viz"
author: "MAX"
date: "2024-09-19"
output: html_document
---

```{r}
library(ggplot2)
library(janitor)
library(tidyverse)
library(USAboundaries)
library(sf)
area <- read_csv("ServiceArea.csv")
HospitalLocations <- read_csv("us_hospital_locations.csv")
library(plotly)
```

```{r}
usmap <- USAboundaries::us_states()
continental_usmap <- usmap %>%
  filter(!name %in% c("Alaska", "Hawaii"))

HospitalLocations_sf <- st_as_sf(HospitalLocations, coords = c("X", "Y"), crs = 4326)
HospitalLocations_sf <- st_transform(HospitalLocations_sf, st_crs(usmap)) 
HospitalLocations_sf <- HospitalLocations_sf %>%  filter(STATE %in% continental_usmap$stusps)

Area2 <- area %>% group_by(StateCode) %>% summarise(n()) 
Area3 <- left_join(x = Area2, y = continental_usmap, by = c("StateCode" = "stusps"))

```


```{r fig.alt= "Hospital Locations and Insurance Plan Coverage in the U.S. From data from the Center for Medicare and Medicaid Services"}

p <- ggplot() +
  geom_sf(data = continental_usmap, fill = "lightgray") + 
  geom_sf(data = Area3, aes(fill = `n()`, geometry = geometry)) +
  scale_fill_gradient(low = "lightyellow", high = "darkblue", name = "Insurance Plans") +  
  geom_sf(data = continental_usmap, color = "white", fill = NA) + 
  geom_point(data = HospitalLocations_sf, aes(x = LONGITUDE, y = LATITUDE), alpha = 0.5, size = 0.6, color = "darkred") +
  theme_minimal() + labs(title = "Hospital Locations and Insurance Plan Coverage in the U.S.", subtitle = "From data from the Center for Medicare and Medicaid Services", x = "Longitude", y = "Latitude")

p

```

Here we have a map of hospital locations and insurance plan coverage in the United States. My main goal in creating this visualization was to gain some context on some of the datasets that my group and I were encountering in our search. Our research question in this case is: What factors make someone more or less likely to receive care? So, of course, I grabbed this hospital locations dataset to take a look and see if Maybe distance could be a problem for a big enough group of people. This dataset also contains information on the county level, so we can even get more precise while trying to take a look at this. Our axes are simple enough, we have latitude and longitude on the x and y respectively. Our red dots are hospitals, and we can see specifically Denver, Colorado, with a big clump of hospitals. Any big city has a big clump of hospitals of course. We also have a color gradient of how many insurance plans were offered by State/County, but more specifically “Health and dental plans offered to individuals and small businesses through the US Health Insurance Marketplace.” Unfortunately, it turns out that this dataset is not quite the best for a country-wide analysis, so I am excited to start narrowing down the search on the map. Some overall takeaways are that clearly our insurance dataset really only has anything to do with The states of Michigan, Texas, and Virginia, so that is one big limitation as well, and also we can see that obviously hospitals are clustered where people are clustered. 
